Research training portfolio

### 2012-11-06 Research paper.tex
RP: wat structuur en flarden toegevoegd
RP: revision, start with classical MT. RTP: start with evaluation

\usepackage{covington}

@@

When one of Starfleet's vessels meets an alien ship in areas where no man has gone before, they can (almost) always communicate with these aliens (almost) instantly; they speak to them in English, and these aliens answer in English. This is possible because, according to the Star Trek lore, the protagonists make use of a so-called Universal Translator. This Universal Translator can (1) recognize speech in every language, (2) translate what it has 'heard' to any other language, (3) make its results audible with speech synthesis and (4) learn itself new languages when needed. Unfortunately, like many technologies in Star Trek, this is far from reality: there have been various attempts at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator. \\\indent
In this paper, I will give a detailed overview of the research project Colibri, which basically is an attempt to improve the technology needed for the second task: automatically translating text from one language to another. More specifically, the project investigates the identification and extraction of constructions in natural language, and how these can be used in Machine Translation. These constructions can be a various lengths, and, importantly, can also contain one or more gaps. The constructions are not identified on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora') with the help of context-sensitive machine-learning techniques.

@@

%Stukje over dat het software is, dus tekst over wat de software inhoudt is tegelijkertijd een beschrijving van wat het softwarepakket inhoudt

@@

The first gap is usually filled with a noun phrase, the second one verb phrase. Of course, not every possible combination of words is a construction; what makes some combinations special? The exact nature of linguistic constructions has been the subject of many publications and even an entire linguistic framework (construction grammar). For this project however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.

@@

\section{Theoretical framework}

@@

% Hier dingen zeggen over hoe er eerst vertaald werd. Of is dat theoretical framework?
% Ook lingusitic relevance: Abstracting fully lexicalised constructions
%Finding semantic subclasses in constructions: from time-expression to time-expression"
%Collapsing constructions with word disjunctions (\he/she/it") or part-of-speech tags
% Correlations with experimental findings
% Switch tasks, cloze tests, reaction times, ...

@@

\section{Achievements so far}
\section{Main outstanding questions}
\section{Strengths and weaknesses}
\section{Possibilities for future research}
\section{Conclusion}
\end{document}

>>>>

\usepackage{covington}
\usepackage{graphicx}

@@

When one of Starfleet's vessels meets an alien ship in areas where no man has gone before, they can (almost) always communicate with these aliens (almost) instantly; they speak to them in English, and these aliens answer in English. This is possible because, according to the Star Trek lore, the protagonists make use of a so-called Universal Translator. This Universal Translator can (1) recognize speech in every language, (2) translate what it has 'heard' to any other language, (3) make its results audible with speech synthesis and (4) learn itself new languages when needed. Unfortunately, like many technologies in Star Trek, this is far from reality: there is lots of active research at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator. \\\indent
In this paper, I will give a detailed overview of the research project Colibri, which basically is an attempt to improve the technology needed for the second task: automatically translating text from one language to another. More specifically, the project investigates the identification and extraction of constructions in natural language, and how these can be used in Machine Translation. These constructions can be of various lengths, and, importantly, can also contain one or more gaps. The constructions are not identified on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora') with the help of context-sensitive machine-learning techniques. The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.

@@

Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology. The following project overview therefore necessarily also is an overview of the software.

@@

The first gap is usually filled with a noun phrase, the second one with a verb phrase. Of course, not every possible combination of words is a construction; what makes some combinations special? The exact nature of linguistic constructions has been the subject of many publications and even an entire linguistic framework (construction grammar). For this project however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.

@@

\section{Theoretical background and framework}

\subsection{Classical Machine Translation}

Classical Machine Translation is best characterized by the use of human created rules. That is, authors of translation algorithms tried to include their knowledge about the languages and translation. There have been four main approaches, summarized in the Vauquois triangle: \\

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{vauquois.png}
\caption{The Vauquois triangle}
\label{fig:vauquois}
\end{figure}

The horizontal arrows represent the approaches, I'll go through them from bottom to top:

\begin{itemize}
\item The direct approach does very little source language analysis, and could thus be viewed as the automatical version of looking up every word in the source language in a dictionary and replacing it with the corresponding word in the target language. This often doesn't yield very good results: if I for example would have translated this paper from my mother tongue Dutch to English word for word, 'would we now not real a good result have'. It should be noted, however, that direct approach can be a little more intelligent than that: %wwat dan?
\item The syntactic approach
\item The semantic approach
\item All of the approaches so far have in common that they require a seperate rule set for each language pair; no or very few rules included for translations from English to German can also be used for translations from English to French. The Interlingua approach is an attempt to solve this problem, by pairing up every language with one general language, the Interlingua. This means a translation from English to German would entail a translation from English to the Interlingua (using the ruleset English $\rightarrow$ Interlingua) and then a translation from the Interlingua to German (using the ruleset Interlingua $\rightarrow$ German). When translating from English to French, the same two steps would be taken, and the same ruleset English $\rightarrow$ Interlingua can be used. This Interlingua is a formal representation.
\end{itemize}

% Voorbeelden?

\subsection{Statistical Machine Translation}

@@

% Ook lingusitic relevance: Abstracting fully lexicalised constructions
%Finding semantic subclasses in constructions: from time-expression to time-expression"
%Collapsing constructions with word disjunctions (\he/she/it") or part-of-speech tags
% Correlations with experimental findings
% Switch tasks, cloze tests, reaction times, ...

@@

\section{Achievements so far}
\section{Main outstanding questions}
\section{Strengths and weaknesses}
\section{Possibilities for future research}
\section{Conclusion}

\section{References}

% Documentatie Colibri
% Maartens thesis

\end{document}

### 2012-11-06 Research training portfolio.tex
Added info table, added research paper outline
Added student number
RP: revision, start with classical MT. RTP: start with evaluation

@@

@@

>>>>

\begin{table}[h]
\begin{tabular}{ll}
Name&Wessel Stoop\\
Student numbers&s0808709 (Nijmegen), u1249664 (Tilburg)\\
Project&Fowlt, a context-based spelling checker\\
Supervisor&Antal van den Bosch\\
Period&September - December 2012\\
\end{tabular}
\end{table}

@@

% Hier moet ergens een beetje beschreven worden hoe zo'n Timblmodule werkt.

\subsubsection{Evaluation}

In this section, I describe the various evaluation measures I used. This is necessary because evaluation of spelling corrector has not been very uniform in the spelling checker literature so far; whereas various scholars only measured how many errors their algorithm can detect in an error list and called it accuracy (e.g. Agirre et al., 1998; Brill \& Moore, 2000; Toutanova \& Moore, 2002; Van Delden et al., 2004), others use or argue for the precision-recall terminology (Reynaert, 2008; Pollock \& Zamora, 1984, although they use other terms for the two concepts) or argue against it (Starlander \& Popescu-Belis, 2002).\\\indent

Although most of the evaluation measures used can be useful and informative, they only tell part of the story. I think it is important to recognize that spelling correction is a multi-step process, and each step comes with its own measures. Using a confusible corrector as part of a spelling corrector, there are three steps: word prediction, error detection and error correction. I'll go through these steps and their evaluation measures one by one.

\paragraph{Word prediction}

\begin{itemize}
\item Accuracy. Although word prediction may be the most complicated part of the confusible corrector, its evaluation is actually the simplest. Calculating rediction accuracy entails removing all confusibles from a text, and see in how many of the cases the language model can correctly predict which word was removed. This percentage can be increased by only counting predictions that are above a certain threshold. Results with multiple thresholds (i.e. to what extent does the prediction improve if the threshold is raised) can be visualized easily with a graph, as shown in the result section.
\end{itemize}

\paragraph{Error detection}
\begin{itemize}
\item The confusion matrix.
\item Recall.
\item Precision.
\item Accuracy.
\end{itemize}

\paragraph{Error correction} Some text.

\paragraph{Problems with precision and recall.} Some text.

@@

\section{References}
% Ene paper Antal
% roc101
% (e.g. Agirre et al., 1998; Brill & Moore, 2000; Toutanova & Moore, 2002; Van Delden et al., 2004)
% (Reynaert, 2008; Pollock & Zamora, 1984, although they use other terminology)
% (Starlander & Popescu-Belis, 2002)
